\bf The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-03-10 paper/forestecology.tex
3060 -> 3000 🏁🏁🏁

\noindent%
{\it Keywords:} forest ecology, interspecific competition, neigbhorhood competition,
tree growth, R, ForestGEO, spatial cross-validation
\vfill

@@

Repeat-censused forest plots offer excellent opportunities to test
neighborhood models of the effect of competition on the growth of trees
(\citet{canham_neighborhood_2004}). Neighborhood models of competition
have been used to: test whether the species identity of a competitor
matters (\citet{uriarte_spatially_2004}); measure species-specific
competition coefficients (\citet{das_effect_2012}
\citet{tatsumi_estimating_2016}); test competing models to see what
structures competitive interactions, e.g.~traits or phylogeny
(\citet{allen_permutation_2020}; \citet{uriarte_trait_2010}); and inform
selective logging practices (\citet{canham_neighborhood_2006}). Although
these are well-described methods, few methods are currently available
for easy application. Here we address this in an R package. We largely
follow the methods presented in \citet{allen_permutation_2020}. The
package is written to model stem radial growth between two censuses
based on neighborhood competition.

\citet{allen_permutation_2020} considers the following model: Let
\(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
species groups. They use the following linear model the average annual
growth in diameter at breast height (DBH) \(y_{ij}\) (in
centimeters/year) of the \(i^{th}\) tree of focal species group \(j\):

\begin{equation}
\label{eq:model}
y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
\end{equation}

where \(\beta_{0,j}\) is the diameter-independent growth rate for group
\(j\); \(\text{dbh}_{ij}\) is the DBH the focal tree at the earlier
census; \(\beta_{\text{dbh},j}\) is the amount of the growth rate
changed depending on diameter for group \(j\); \(\text{BA}_{ijk}\) is
the sum of the basal area of all trees of competitor species group
\(k\); \(\lambda_{jk}\) is the change in growth for individuals of group
\(j\) from nearby competitors of group \(k\); and \(\epsilon_{ij}\) is a
random error term distributed \(\text{Normal}(0, \sigma^2)\). They
estimate all parameters via Bayesian linear regression while exploiting
Normal/Inverse Gamma conjugacy to derive closed-form solutions to all
posterior distributions\footnote{See S1 Appendix of
  \citet{allen_permutation_2020}, available at
  \url{https://doi.org/10.1371/journal.pone.0229930.s004}}. These
closed-form solutions for the posterior distributions are in contrast to
approximations of all posteriors via computationally expensive Markov
Chain Monte Carlo algorithms.

In order to evaluate whether competitor species identity matters,
\citet{allen_permutation_2020} run a permutation test where a null
hypothesis of no species grouping-specific effects of competition is
assumed, and thus the species identity of all competitors can be
permuted/shuffled:

\begin{eqnarray}
\label{eq:permutation-hypothesis-test}
&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K\\
\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different}
\end{eqnarray}

Furthermore, in order to account for the spatial autocorrelation
inherent to forest data in their estimates of out-of-sample model error,
\citet{allen_permutation_2020} use spatial cross-validation. Estimates
of model error that do not account for this spatial dependency tend to
underestimate the true model error
\citep{roberts_cross-validation_2017}.

We introduce the \texttt{forestecology} R package providing methods and
data for forest ecology model fitting and assessment, available on CRAN
(\url{https://cran.r-project.org/web/packages/forestecology/index.html})
and on GitHub (\url{https://github.com/rudeboybert/forestecology}). The
package implements all aspects of the model in Equation \ref{eq:model}:
model fitting and generating predicted values, evaluating the effect of
competitor species identity using permutation tests, and evaluating
model performance using spatial cross-validation.

The package designed with ``tidy'' design principles in mind
\citep{wickham_welcome_2019}. Much like all \texttt{tidyverse} packages,
\texttt{forestecology} is has verb-named functions that can be modularly
composed using the pipe \texttt{\%\textgreater{}\%} operator to complete
all the necessary steps in the analysis sequence
\citep{bache_pipe_2020}. Furthermore, the inputs and outputs of most of
our functions use the same ``simple features for R'' data structures
from the \texttt{sf} package for standardized and
\texttt{tidyverse}-friendly wrangling and visualizing of spatial data
\citep{pebesma_simple_2018}

Currently the package only implements the Bayesian linear regression
model detailed in Equation \ref{eq:model}. As we demonstrate in Section
\ref{model-fit-predict} however, the fitting of this model is
self-contained in a single function \texttt{comp\_bayes\_lm()} which
returns an object of S3 class type \texttt{comp\_bayes\_lm}. This class
has generic methods implemented to print, make predictions using, and
plot all results. Therefore the package can be modularly extended to fit
other models as long as they are coded similarly as
\texttt{comp\_bayes\_lm()} and have equivalent generic methods
implemented.

@@

We present a case-study of the \texttt{forestecology} package's
functionality on data from the Smithsonian Conservation Biology
Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA,
part of the ForestGEO global network of research sites
\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}
\citep{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
the intersection of three of the major physiographic provinces of the
eastern US---the Blue Ridge, Ridge and Valley, and Piedmont
provinces---and is adjacent to the northern end of Shenandoah National
Park.

The package has the following ecological goals: to evaluate i) the
effect of competitor species identity using permutation tests and ii)
model performance using spatial cross-validation. We outline a basic
analysis sequence comprising of the following four main steps:

@@

The first step is to compute the growth of trees using data from two
censuses. \texttt{compute\_growth()} computes the average annual growth
based on census data that roughly follows ForestGEO standards. Despite
such standards, minor variations will still exist between sites, thereby
necessitating some data wrangling. For example, the SCBI site records
all DBH values in millimeters \citep{bourg_initial_2013}, whereas the
Michigan Big Woods site records them in centimeters
\citep{allen_michigan_2020}.

We first load both 2008 and 2014 SCBI census \texttt{.csv} files as they
existed on GitHub on 2020/11/20 and perform some data wrangling to both
data sets \citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}.
We then only consider a 9 ha subsection of the 25.6 ha of the SCBI site
in order to speed up computation for this example: \texttt{gx} from
0--300 instead of 0--400 and \texttt{gy} from 300--600 instead of
0--640.

@@

These two data frames are then used as inputs to
\texttt{compute\_growth()}, along with the \texttt{id} argument
specifying the variable that uniquely identifies each tree-stem. Note
that we also discard all resprouts in the later census with
\texttt{code\ ==\ R}, since we are only interested in the growth of
surviving, and not resprouted, stems.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}
\NormalTok{growth_scbi}
\CommentTok{## Simple feature collection with 7954 features and 8 fields}
\CommentTok{## geometry type:  POINT}
\CommentTok{## dimension:      XY}
\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
\CommentTok{## CRS:            NA}
\CommentTok{## # A tibble: 7,954 x 9}
\CommentTok{##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth}
\CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>}
\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103}
\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150}
\CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140}
\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141}
\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161}
\CommentTok{## # ... with 7,949 more rows, and 1 more variable: geometry <POINT>}
\end{Highlighting}
\end{Shaded}

The output \texttt{growth\_scbi} is a data frame of class \texttt{sf}
that includes variables the average annual \texttt{growth} in DBH (cm
\(\cdot\) y\textsuperscript{-1}) for all stems that were alive at both
time points, the \texttt{sf} package's encoding of geolocations of
\texttt{geometry} type \texttt{\textless{}POINT\textgreater{}}, and the
species variable \texttt{sp} converted to a factor. Given that
\texttt{growth\_scbi} is of class \texttt{sf}, it can be easily plotted
in \texttt{ggplot2} using \texttt{geom\_sf()} as seen in Figure
\ref{fig:scbi-trees}.

@@

\caption{Compute growth of trees based on census data: Map with growth of a random sample of 500 trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot.}\label{fig:scbi-trees}
\end{figure}

@@

The next step is to add spatial information to \texttt{growth\_scbi}.
The first element we add is a ``buffer region'' to the periphery of the
study region. Since some of our model's explanatory variables are
cumulative, we must ensure that all trees being modeled are not biased
to have different neighbor structures. This is of concern for trees at
the boundary of the study region whose neighbors will not all be
included in the censused stems. To account for such edge effects only
trees that are not part of this buffer region, i.e.~are part of the
interior of the study region, will have their growth modeled
\citep{waller_applied_2004}.

Our model of interspecific competition relies on a spatial definition of
who competitor trees are: all trees within a distance
\texttt{comp\_dist} of a focal tree. In our case we set
\texttt{comp\_dist} to 7.5m, a value informed by other studies
\citep[\citet{uriarte_spatially_2004},
\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We use
\texttt{comp\_dist} and a manually constructed \texttt{sf}
representation of the study region's boundary as inputs to
\texttt{add\_buffer\_variable()} to add a \texttt{buffer} boolean
variable to \texttt{growth\_scbi}. All trees with \texttt{buffer} equal
to \texttt{FALSE} will be our focal trees whose growth will be modeled,
whereas those with \texttt{TRUE} will only act as competitor trees.

@@

The second element of spatial information we add are blocks
corresponding to folds of a spatial cross-validation algorithm.
Conventional cross-validation algorithms assign observations to folds by
randomly resampling individual observations that are assumed
independent. In the case of forest census data however, observations
exhibit spatial autocorrelation. We therefore incorporate this
dependence into the cross-validation algorithm by resampling spatial
blocks of trees \citep[
\citet{pohjankukka_estimating_2017}]{roberts_cross-validation_2017}.

We first manually an \texttt{sf} object defining four folds that
partition the study region. We then use the output of the
\texttt{spatialBlock()} function from the \texttt{blockCV} package to
associate each tree in \texttt{growth\_scbi} to the correct
\texttt{foldID} \citep{valavi_blockcv_2019}.\footnote{In the Supporting
  Information we present an example where the folds themselves are
  created automatically, as opposed to manually as in the example.} This
\texttt{foldID} variable will be used in Section
\ref{spatial-cross-validation}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Manually define spatial blocks to act as folds}
\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{))}
\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{))}
\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}

@@

\CommentTok{# Associate each observation to a fold}
\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }
  \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
\NormalTok{)}

@@

\caption{Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4). The location of each tree is marked with an integer indicating its fold, with folds delineated with solid lines. The color of each digit indicates whether the tree is part of the buffer region (and thus will only be considered as a competitor tree in our model) or is part of the interior of the study region (and thus is a focal tree whose growth is of modeled interest).}\label{fig:scbi-spatial-information}
\end{figure}

@@

The next step is to identify all focal trees and their corresponding
competitor trees. More specifically, identify all trees that are not
part of the buffer region, have a valid \texttt{growth} measurement, and
have at least one neighbor within 7.5m. We do this using
\texttt{create\_focal\_vs\_comp()}, which takes the previously detailed
arguments \texttt{comp\_dist} and \texttt{id} as well as the \texttt{sf}
representation of the spatial cross-validation blocks and returns a new
data frame \texttt{focal\_vs\_comp\_scbi}.

@@

We observe 4 variables describing 20 competitor trees: their unique
tree-stem ID, their distance to the focal tree (all \(\leq\) 7.5), their
species, and their basal area (in m\(^2\)) calculated as
\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) for \(DBH\) in cm from the
earlier census. Saving competitor information in list-column format
minimizes redundancy since we do not need to repeat information on the
focal tree 20 times. We visualize the spatial distribution of these
trees in Figure \ref{fig:scbi-focal-vs-comp-map}.

@@

\caption{Identify all focal and corresponding competitor trees: The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle.}\label{fig:scbi-focal-vs-comp-map}
\end{figure}

@@

The final step is to fit the competition Bayesian linear regression
model for tree growth outlined in Equation \ref{eq:model} using
\texttt{comp\_bayes\_lm()}. This function has an option to specify prior
distributions on all parameters of interest, chosen here to be the
defaults detailed in \texttt{?comp\_bayes\_lm}.

@@

The resutling \texttt{comp\_bayes\_lm\_scbi} is an object of S3 class
type \texttt{comp\_bayes\_lm} containing the posterior values of all
parameters of our model. Furthermore, this class of object includes
generics for three methods. First, the generic for \texttt{print()}
displays the names of all prior and posterior parameters along with the
model formula:

@@

Next, the generic for \texttt{predict()} takes the posterior parameter
values in \texttt{comp\_bayes\_lm\_scbi} and a \texttt{newdata} data
frame and outputs a vector \texttt{growth\_hat} of predicted values
\(\widehat{y}\) of the DBH for each focal tree computed from the
posterior predictive distribution.

@@

We then compare the observed and predicted growths to compute the root
mean squared error (RMSE) of our model fit.

@@

\caption{Fit model: Posterior distributions of all parameters. For compactness we include only three species.}\label{fig:scbi-posterior-viz}
\end{figure}

>>>>

\noindent%
{\it Keywords:} forest ecology, interspecific competition, neighborhood competition,
tree growth, R, ForestGEO, spatial cross-validation
\vfill

@@

Repeat-censused forest plots offer excellent opportunities to test
neighborhood models of the effect of competition on the growth of trees
(\citet{canham_neighborhood_2004}). Neighborhood models of competition
have been used to: test whether the species identity of a competitor
matters (\citet{uriarte_spatially_2004}); measure species-specific
competition coefficients (\citet{das_effect_2012}
\citet{tatsumi_estimating_2016}); test competing models to see what
structures competitive interactions, e.g.~traits or phylogeny
(\citet{allen_permutation_2020}; \citet{uriarte_trait_2010}); and inform
selective logging practices (\citet{canham_neighborhood_2006}). Although
these are well-described methods, few methods are currently available
for easy application.

We address this shortcoming with the \texttt{forestecology} R package
providing methods and data for forest ecology model fitting and
assessment, available on CRAN
(\url{https://cran.r-project.org/web/packages/forestecology/index.html})
and on GitHub (\url{https://github.com/rudeboybert/forestecology}). The
package is written to model stem radial growth between two censuses
based on neighborhood competition, largely following the methods in
\citet{allen_permutation_2020}.

Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
species groups. The average annual growth in diameter at breast height
(DBH) \(y_{ij}\) (in centimeters/year) of the \(i^{th}\) tree of focal
species group \(j\) is modeled as

\begin{equation}
\label{eq:model}
y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
\end{equation}

where \(\beta_{0,j}\) is the diameter-independent growth rate for group
\(j\); \(\text{dbh}_{ij}\) is the DBH of the focal tree at the earlier
census for group \(j\) and \(\beta_{\text{dbh},j}\) quantifies its
corresponding growth rate change; \(\text{BA}_{ijk}\) is the sum of the
basal area of all trees of competitor species group \(k\) and
\(\lambda_{jk}\) quantifies the corresponding change in growth for
individuals of group \(j\) from these competitors; and \(\epsilon_{ij}\)
is a random error term distributed \(\text{Normal}(0, \sigma^2)\). They
estimate all parameters via Bayesian linear regression, while exploiting
Normal/Inverse Gamma conjugacy to derive closed-form solutions to all
posterior distributions\footnote{See S1 Appendix of
  \citet{allen_permutation_2020}, available at
  \url{https://doi.org/10.1371/journal.pone.0229930.s004}}. These
closed-form solutions are not as computationally expensive as
approximations from Markov Chain Monte Carlo algorithms.

To evaluate whether competitor species identity matters,
\citet{allen_permutation_2020} run a permutation test where a null
hypothesis of no species grouping-specific effects of competition is
assumed, thus the species identity of all competitors can be permuted:

\begin{eqnarray}
\label{eq:permutation-hypothesis-test}
&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K\\
\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different} \nonumber
\end{eqnarray}

Furthermore, to account for the spatial autocorrelation in their
estimates of out-of-sample model error, \citet{allen_permutation_2020}
use spatial cross-validation. Estimates of model error that do not
account for this dependence tend to underestimate the true model error
\citep{roberts_cross-validation_2017}.

The package is designed with ``tidy'' design principles in mind
\citep{wickham_welcome_2019}. Much like all \texttt{tidyverse} packages,
\texttt{forestecology} has verb-named functions that can be modularly
composed using the pipe \texttt{\%\textgreater{}\%} operator to
sequentially complete all necessary analysis steps
\citep{bache_pipe_2020}. Furthermore, the inputs and outputs of most
functions use the same ``simple features for R'' data structures from
the \texttt{sf} package, a package for standardized and
\texttt{tidyverse}-friendly wrangling and visualizing of spatial data
\citep{pebesma_simple_2018}.

Currently the package only implements the Bayesian linear regression
model detailed in Equation \ref{eq:model}. As we demonstrate in Section
\ref{model-fit-predict} however, the fitting of this model is
self-contained in a single function \texttt{comp\_bayes\_lm()} which
returns an object of S3 class type \texttt{comp\_bayes\_lm}. This class
has generic methods implemented to print, make predictions, and plot all
results. Therefore the package can be modularly extended to fit other
models as long as they are coded similarly to \texttt{comp\_bayes\_lm()}
and have equivalent generic methods implemented.

@@

We present a case-study of \texttt{forestecology}'s functionality on
data from the Smithsonian Conservation Biology Institute (SCBI) large
forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO
global network of research sites
\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}
\citep{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
the intersection of three of the major physiographic provinces of the
eastern US---the Blue Ridge, Ridge and Valley, and Piedmont
provinces---and is adjacent to the northern end of Shenandoah National
Park.

The package has the following goals: to evaluate i) the effect of
competitor species identity using permutation tests and ii) model
performance using spatial cross-validation. We outline the four-step
basic analysis sequence:

@@

We first compute the growth of trees using data from two censuses.
\texttt{compute\_growth()} computes the average annual growth based on
census data that roughly follows ForestGEO standards. Despite such
standards, minor variations will still exist between sites, thereby
necessitating some data wrangling. For example, the SCBI site records
all DBH values in millimeters \citep{bourg_initial_2013}, whereas the
Michigan Big Woods site used in \citet{allen_permutation_2020} records
them in centimeters \citep{allen_michigan_2020}.

We load both 2008 and 2014 SCBI census \texttt{.csv} files as they
existed on GitHub on 2020/11/20 and perform minor data wrangling
\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. We then
only consider a 9 ha subsection of the 25.6 ha of the site to speed up
computation for this example: \texttt{gx} from 0--300 instead of 0--400
and \texttt{gy} from 300--600 instead of 0--640.

@@

These two data frames are then used as inputs to
\texttt{compute\_growth()}, along with \texttt{id} specifying the
variable that uniquely identifies each tree-stem. We also discard all
resprouts with \texttt{code\ ==\ R} in the later census, since we are
only interested in the growth of surviving, and not resprouted, stems.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}
\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, dbh1, dbh2, growth, geometry)}
\CommentTok{## Simple feature collection with 7954 features and 5 fields}
\CommentTok{## geometry type:  POINT}
\CommentTok{## dimension:      XY}
\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
\CommentTok{## CRS:            NA}
\CommentTok{## # A tibble: 7,954 x 6}
\CommentTok{##   stemID sp     dbh1  dbh2 growth   geometry}
\CommentTok{##    <dbl> <fct> <dbl> <dbl>  <dbl>    <POINT>}
\CommentTok{## 1      4 nysy  13.6   14.2  0.103 (14.2 428)}
\CommentTok{## 2      5 havi   8.8    9.6  0.150  (9.4 436)}
\CommentTok{## 3      6 havi   3.25   4    0.140  (1.3 434)}
\CommentTok{## 4     77 qual  65.2   66    0.141 (34.7 307)}
\CommentTok{## 5     79 tiam  47.7   46.8 -0.161   (40 381)}
\CommentTok{## # ... with 7,949 more rows}
\end{Highlighting}
\end{Shaded}

The output \texttt{growth\_scbi} is a data frame of class \texttt{sf}
that includes among other variables the species variable \texttt{sp}
converted to a factor, the average annual \texttt{growth} in DBH (cm
\(\cdot\) y\textsuperscript{-1}) for all stems that were alive at both
time points, and the \texttt{sf} package's encoding of geolocations of
\texttt{geometry} type \texttt{\textless{}POINT\textgreater{}}. Given
that \texttt{growth\_scbi} is of class \texttt{sf}, it can be easily
plotted in \texttt{ggplot2} using \texttt{geom\_sf()} as seen in Figure
\ref{fig:scbi-trees}.

@@

\caption{Compute growth of trees based on census data. A map of the growth of a random sample of 500 trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot.}\label{fig:scbi-trees}
\end{figure}

@@

We then add spatial information to \texttt{growth\_scbi}. We first add a
``buffer region'' to the periphery of the study region. Since some of
our model's explanatory variables are cumulative, we must ensure that
all trees being modeled are not biased to have different neighbor
structures. This is of concern for trees at the boundary of the study
region who will not have all their neighbors included in the census
stems. To account for such edge effects, only trees that are not part of
this buffer region, i.e.~are part of the interior of the study region,
will have their growth modeled \citep{waller_applied_2004}.

Our model of interspecific competition relies on a spatial definition of
who competitor trees are: all trees within a distance
\texttt{comp\_dist} of a focal tree. We set \texttt{comp\_dist} to 7.5m,
a value informed by other studies \citep[\citet{uriarte_spatially_2004},
\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We use
\texttt{comp\_dist} and a manually constructed \texttt{sf}
representation of the study region's boundary as inputs to
\texttt{add\_buffer\_variable()} to add a \texttt{buffer} boolean
variable to \texttt{growth\_scbi}. All trees with \texttt{buffer} equal
to \texttt{FALSE} will be our focal trees whose growth will be modeled,
whereas those with \texttt{TRUE} will only act as competitor trees.

@@

The second element of spatial information we add are blocks
corresponding to folds of a spatial cross-validation algorithm.
Conventional cross-validation algorithms assign individual observations
to folds by randomly resampling them all while assuming they are
statistically independent. In the case of forest census data however,
observations exhibit spatial autocorrelation. We therefore incorporate
this dependence into the cross-validation algorithm by resampling
spatial blocks of trees \citep[
\citet{pohjankukka_estimating_2017}]{roberts_cross-validation_2017}.

We first manually define an \texttt{sf} object defining four folds that
partition the study region. We then use the output of the
\texttt{spatialBlock()} function from the \texttt{blockCV} package to
associate each tree in \texttt{growth\_scbi} to the correct
\texttt{foldID} \citep{valavi_blockcv_2019}.\footnote{In the Supporting
  Information we present an example where the folds themselves are
  created automatically, as opposed to manually as in the example.} This
\texttt{foldID} variable will be used in Section
\ref{spatial-cross-validation}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Manually define spatial blocks to act as folds}
\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}

@@

\CommentTok{# Associate each observation to a fold}
\NormalTok{spatial_block_scbi <-}\StringTok{ }
\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}

@@

\caption{Add spatial information. A buffer region and spatial cross-validation blocks 1 througth 4. The location of each tree is marked with its fold number where the folds are delineated with solid lines. The color of each digit indicates whether the tree is part of the buffer region (thus will only be considered as a competitor tree) or is part of the interior of the study region (thus is a focal tree whose growth is of modeled interest).}\label{fig:scbi-spatial-information}
\end{figure}

@@

We then identify all focal trees and their corresponding competitor
trees. More specifically, identify all trees that are not part of the
buffer region, have a valid \texttt{growth} measurement, and have at
least one neighbor within 7.5m. We do this using
\texttt{create\_focal\_vs\_comp()}, which takes the previously detailed
\texttt{comp\_dist} and \texttt{id} arguments as well as the \texttt{sf}
representation of the spatial cross-validation blocks and returns a new
data frame \texttt{focal\_vs\_comp\_scbi}.

@@

We observe 4 variables describing 20 competitor trees: the unique
tree-stem ID, the distance to the focal tree (all \(\leq\) 7.5), the
species, and the basal area (in m\(^2\)) calculated as
\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) for the DBH in cm from the
earlier census. Saving competitor information in list-column format
minimizes redundancy since we do not need to repeat information on the
focal tree 20 times. We visualize the spatial distribution of these
trees in Figure \ref{fig:scbi-focal-vs-comp-map}.

@@

\caption{Identify all focal and corresponding competitor trees. The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle.}\label{fig:scbi-focal-vs-comp-map}
\end{figure}

@@

Lastly, we fit the competition Bayesian linear regression model for tree
growth outlined in Equation \ref{eq:model} using
\texttt{comp\_bayes\_lm()}. This function has an option to specify prior
distributions of all parameters, chosen here to be the defaults detailed
in \texttt{?comp\_bayes\_lm}.

@@

The resulting\texttt{comp\_bayes\_lm\_scbi} is an object of S3 class
type \texttt{comp\_bayes\_lm} containing the posterior values of all
parameters. Furthermore, this class includes generics for three methods.
First, the generic for \texttt{print()} displays the names of all prior
and posterior parameters and the model formula:

@@

Next, the generic for \texttt{predict()} takes the posterior parameter
values in \texttt{comp\_bayes\_lm\_scbi} and a \texttt{newdata} data
frame, and outputs a vector \texttt{growth\_hat} of predicted DBH values
\(\widehat{y_{ij}}\) computed from the posterior predictive
distribution.

@@

We can now compare the observed and predicted growths to compute the
root mean squared error (RMSE) of our model:

@@

\caption{Fit model. Posterior distributions of all parameters. For compactness we include only three species.}\label{fig:scbi-posterior-viz}
\end{figure}

For many users the visualizations of \(\lambda_{j,k}\) will be of
particular interest as they provide insight into species competitive
interactions, where negative values indicate a competitor species which
slows the growth of a focal species. Here, for example, we see that
tulip poplars (litu) have a strong negative effect on the growth of
conspecifics but relatively lesser effect on pignut hickory (cagl) and
red oak (quru) neighbors.
